### 2021-05-13 assertions/doc/thesis-bernhardt/chapters/5-offline_design_tezos.tex
Final revision part 1

As a preliminary for the compiler implementation, \secref{sec:ext_michelson} identifies necessary extensions to Michelson and Tezos' VM, in order to facilitate the distributed assertion scheme. Although the extensions are formalized in this thesis, implementing the extension is part of the protocol amendment. Furthermore, the compiler needs to generate the target code according to an on-chain orchestration scheme between the assertion and parent contract. \secref{sec:orchestration} explores possible strategies for such a scheme, and discusses their strengths and weaknesses. Based on a resulting proposal for the orchestration, \secref{sec:cost_analysis_distributed} performs a cost analysis of the approach of distributed assertion checking and evaluates if, and under which conditions, it can increase the scalability in blockchain applications.

As a foundation for the other sections, \secref{sec:tezos} provides important background on the Tezos blockchain, its built-in Smart Contract language Michelson and some relevant tools and projects within its ecosystem.

@@

In addition to user accounts associated with a public key, Tezos supports Smart Contracts, which are written in the built-in language Michelson. Tezos' transaction fee system is similar to Ethereum's \cite{wood_ethereum_2021} - it is gas-instrumented and besides a base fee, every operation and byte of storage during contract execution has to be paid for by the user. However, Tezos imposes a hard cap on the amount of gas that can be consumed per operation (including internal transactions) \cite{tezos_docs}\cite{morley_repo}, whereas Ethereum limits the gas quota only in respect to blocks \cite{wood_ethereum_2021}.\\
Tezos is written in the multi-paradigm programming language OCaml. Compiling its source code yields five essential binaries: the  node, baker, endorser, accuser and client. The node is the entity connecting to the peer-to-peer network and keeping a copy of the chain. Bakers are responsible for producing new blocks, the endorsers for validating and signing new blocks and the accusers to call out bakers or endorsers which double-sign or -endorse. The client provides a command line interface to interact with nodes through remote procedure calls (RPC).

\subsection{Proof-of-stake in Tezos}
In Tezos, contracts which have staked a minimum amount of tokens (called a roll), can participate in the consensus mechanism. When participating, a contract has a chance to obtain the role of either a baker or an endorser. Contracts not owning enough tokens or infrastructure, to participate directly, can delegate their baking and endorsing rights to other contracts. The rights to bake or endorse are determined and assigned at the beginning of each cycle, which consists of a specified number of blocks. For baking, a random roll is selected for each block level and the rights are assigned to its owner. The block produced by that baker is signed by a fixed number of endorsers\footnote{For consistency, endorsers are referred to as ``validators'' in the remainder of this chapter} (32 as of protocol 007 Delphi), which also have been assigned endorsing rights for this block level by a random selection of rolls. Since participants can stake more than one roll, they may be assigned several endorsement slots at the same block level.

@@

The concrete syntax of Michelson is called Micheline. A program is represented as Micheline nodes, which can be one of the following constructs:
\begin{itemize}
\item A constant of type integer (in decimal notation)
\item A constant of type string
\item A byte sequence in hexadecimal notation
\item An application of a language primitive to a sequence of nodes
\item A sequence of nodes
\end{itemize}
For documentation, readability and additional type constraints, Michelson and Micheline also offer three types of annotations - type, variable and field or constructor annotations, which are labelled with a unique special character in Micheline. The toplevel structure of a Smart Contract consists of a sequence of the three primitives \texttt{parameter, storage} and \texttt{code}. They declare the type of the input parameter, the storage type and the script, which is interpreted when the contract is invoked. The full grammar of Micheline and Michelson can be found in the Tezos developer resources \cite{tezos_docs}.

\subsubsection{Entrypoints}
Unlike Ethereum's contract language Solidity, Michelson doesn't have a concept of named functions with a dedicated parameter type. Instead of named functions, Michelson contracts model separate entrypoints by taking a sum type as an input parameter. The type constructors can optionally be tagged with a field annotation, which can be considered  as an entrypoint name. The sum type is built by nesting the \texttt{or} type, which has the constructors \texttt{left} and \texttt{right}.

As an example, the contract in \lstref{lst:entrypoints} has two entrypoints --- entrypoint \texttt{A}, which expects an integer wrapped in the \texttt{left} constructor, e.g. \texttt{left 1}, and entrypoint \texttt{B}, which expects a string wrapped in the \texttt{right} constructor, e.g. \texttt{right "hello world"}.
\begin{lstlisting}[language=Michelson, numbers=none, caption=Michelson contract with two entrypoints, label=lst:entrypoints]
parameter (or (int %A) (string %B));
storage unit;
code {
  ...
  IF_LEFT { ... (* entrypoint A *)}
          { ... (* entrypoint B *)}
}
\end{lstlisting}
By adding the constructor annotations \texttt{A} and \texttt{B}, the entrypoints can be invoked explicitly. The passed parameter value is then wrapped automatically into the respective constructor. Furthermore, it's possible to declare a \texttt{default} entrypoint, which is invoked when no explicit tag is specified in the transaction. By default, the default entrypoint is assigned to the root of the parameter type.

\subsubsection{Internal operations}\label{sec:internal_ops}
Smart Contracts are able to invoke other Smart Contracts by emitting internal operations. The return type of Michelson programs is a tuple of a list of internal operations and the updated storage. After a contract completes, the operations in the returned list are run in sequence. Since internal operations may in turn also emit internal operations, the interpreter puts all emitted operations into a queue, which is processed in order. As an example, \figref{fig:internal_ops} shows an execution order of two external transactions and their internal operations in sequence.
\begin{figure}[h]
\centering
\includegraphics[width=0.5\linewidth]{figures/5-offline_tezos/internal_ops}
\captionsource{Execution order of two external operations and their internal operations}{\cite{tezos_docs}}
\label{fig:internal_ops}
\end{figure}

@@

\subsection{Gas model}
Tezos stores and transmits data as byte sequences --- in order to obtain a typed representation of their value for interpretation, a byte sequence is first deserialised into Micheline and subsequently parsed into a typed AST. Conversely, before transmitting or storing something on the blockchain, the typed representation is first unparsed into Micheline and then serialised. Since these working steps also entail computational effort, they consume gas and thus account for some of the transaction fees. The gas consumption of an operation (e.g., a transaction or a contract origination) is thus composed of the following positions \cite{morley_gasmodel}\cite{tezos_repo} on top of a base gas fee:
\begin{enumerate}
\item \textbf{Reading costs} --- apply for reading the contract code and storage from the blockchain and depend on the amount of bytes read
\item \textbf{Deserialization costs} --- depend on the Micheline expression and the size of the byte sequence. Apply lazily, thus the deserialization of each value only has to be paid for once
\item \textbf{Parsing costs} --- are composed of three types of costs:
	\begin{itemize}
	\item type parsing --- apply when a Micheline node is converted into a valid type
	\item data parsing --- apply when a node is converted into the value of a known type and depends on the size of the value
	\item code parsing --- apply when the instructions of a program are type checked and depends on the number and types of instructions. Some instructions are expensive, such as \texttt{IF} (both branches need to be checked and compared) or \texttt{CREATE\_CONTRACT} (requires type check of the invoked contract)
	\end{itemize}
\item \textbf{Type comparison costs} --- apply when types are checked for equality (e.g. when type checking the branches of the \texttt{IF} instruction) or during contract interpretation
\item \textbf{Interpretation costs} --- apply for each interpreted instruction and depend on their complexity.
\item \textbf{Unparsing costs} --- apply when converting typed data to untyped Micheline nodes and depend on the data type and size
\item \textbf{Serialization costs} --- apply when serializing Micheline nodes to byte sequences and depend on the expression and its size
\item \textbf{Writing costs} --- apply when writing to the internal database and depend on the amount of bytes. A base gas fee is burned for each additional byte of storage.
\end{enumerate}
If a transaction emits internal operations, their gas consumption is computed in the same way and is added to the total gas consumption of the external operation.

\subsection{Developer tools in the Tezos ecosystem}
This section describes some tools and projects within Tezos' ecosystem, which are relevant to the development of the toolchain or the protocol amendment.

@@

\subsubsection{Testing tools}
Before a new protocol is proposed to the Tezos network, it has to be thoroughly tested with system, integration and regression tests. This requires a sandboxed network to simulate the real peer-to-peer network. Tezos' development environment provides the two testing frameworks Flextesa (Flexible network sandboxes)\cite{tezos_docs} and its successor Tezt \cite{tezos_docs}. They allow to configure and run a small, fully functional sandboxed test network including nodes, bakers, endorsers and accusers. Besides testing new protocols, they can also be used to interactively test Smart Contracts.

\subsubsection{High-level languages}\label{sec:languages}
Michelson is a compilation target for various high-level languages that provide a more user-friendly and intuitive way of writing Smart Contracts. Additionally, some of them provide development environments, testing or verification tools. The following list introduces three prominent languages, which compile to Michelson:
\begin{description}
\item[\textbf{Liquidity}] \cite{liquidity} is a language with an OCaml-like syntax, which allows to express contracts in a functional way. It supports using local variables instead of stack manipulations. Its module system can be used to write reusable contract code or libraries. Besides an optimizing compiler, the project also includes a decompiler to compile Michelson programs to Liquidity source code.
\item[\textbf{SmartPy}] \cite{smartpy} is a language available through a Python library and lets developers write contracts and tests using Python syntax and structures (such as classes). Its developer suite includes, i.a., a compiler, a simulation engine for testing contracts and an online editor.
\item [\textbf{Ligo}] \cite{ligo} offers multiple syntax flavours close to Pascal, OCaml or Reason. Besides compilation, the toolchain allows to invoke or evaluate the generated code locally. The project provides Ligo as OCaml packages, which can be integrated into other projects.
\end{description}

@@

\subsection{Language semantics and type system}\label{sec:michelson_semantics}
Michelson is interpreted purely functional; the interpreter takes the current stack and an operation and builds a return stack from the initial one, without causing any side effects. The definition of the recursive interpreter is given in the form of a list of rules, comprising of all possible inputs (i.e., program and stack types) and the respective output stack type of the computation. Each rule is of the following form: 
\begin{lstlisting}[caption=Selection rules in the Michelson interpreter \cite{tezos_docs}, language=, numbers=none, label=lst:rules]
> (syntax pattern) / (initial stack pattern)  =>  (result stack pattern)
    iff (conditions)
    where (recursions)
    and (more recursions)
\end{lstlisting}
For each valid program and initial stack, exactly one rule applies. Following the keyword \texttt{iff}, the rule can add extra conditions over values on the stack. If the result depends on the results of other program interpretations, such as conditionals or function calls, the rule can contain recursive rules in the \texttt{where} or \texttt{and} clauses. These clauses represent a recursive interpretation of an intermediate program. The rule thus only applies, if the interpretation of the intermediate result matches the expected pattern.

@@

\subsection{random}\label{sec:random}
Since Smart Contracts generally need to be deterministic, s.t. the network can reach a consensus about the state of the ledger \cite{chatterjee_probabilistic_2019}, the prevalent workarounds for generating pseudo-random numbers cannot be used to implement the random generators in assertion contracts. The scheme explicitly requires the validators to check distinct elements in the search space. Common schemes for obtaining random values, like oracles or using block attributes as seeds \cite{chatterjee_probabilistic_2019}, would generate the same values for all validators. Thus, the VM needs to provide an instruction to generate a ``real'' random value exclusively for assertion checking. The instruction should not be available in the normal execution mode, in order to preserve the deterministic behaviour of normal Smart Contracts.

\lstref{lst:rand_type} specifies the typing and selection rule of a new instruction \texttt{random} for integers (such an instruction can also be provided for other data types, such as \texttt{nat}, \texttt{string} or \texttt{mutez}). It consumes an offset and a positive range from the stack and pushes a randomly generated integer to the top. The typing rule (line 1) specifies, that the offset should be of type \texttt{integer}, and the range of type \texttt{nat} (i.e., a natural number).
\lstset{upquote=true}
\begin{lstlisting}[caption=Typing and selection rules of the integer \texttt{random} instruction, language=, showlines=true, label=lst:rand_type]
:: int : nat: 'A -> int : 'A
> RANDOM / offset : range : S  => int : S
\end{lstlisting}

Within the Michelson interpreter, the selection of rules is implemented as a huge match case \cite{tezos_repo}. If the rule for \texttt{random} applies, the instruction could be computed in OCaml as indicated in \lstref{lst:rand_impl}.
\begin{lstlisting}[caption=Simplified evaluation of \texttt{random} in the Michelson interpreter, language=, label=lst:rand_impl]
match (instruction, stack) with
...
| (Random (offset, (range, rest))) ->
   Random.self_init;
   let rand_int = offset + Random.int(range) in
   return (rand_int, rest)  (* resulting stack *)
\end{lstlisting}

\subsection{nth}\label{sec:nth}
Michelson currently does not provide any predefined indexed access operators for data types like lists, strings or bytes. Accessing a random element of a sequence can, of course, be implemented using low-level instructions. However, because such operations are used frequently in many of the given use-cases, adding it as a predefined primitive for some data types might be justifiable. It reduces the required code to express the indexing, and thus not only improves the readability of the contract code, but also lowers the costs of the contract origination and execution.

\lstref{lst:nth_type} specifies the typing and selection rules of an instruction \texttt{nth} to access the $n^{th}$ element of a list. It consumes a list with elements of type \texttt{'a}, and an index of type \texttt{nat} from the top of the stack. After the computation, on top of the resulting stack is an optional value of the element type \texttt{'a}. A selection rule (line 2 and 3) is given for each of the two possible result stack patterns, i.e., a value of type \texttt{'a} or an undefined value if the given index is out of bounds.
\begin{lstlisting}[caption=Selection and type rule of the list \texttt{nth} instruction, language=, label=lst:nth_type]
:: (list 'a) : nat : 'A -> option 'a : 'A
> NTH / l : index : S  =>  Some 'a: S
     iff index is within bounds
> NTH / l : index : S  =>  None 'a: S
     iff index is out of bounds
\end{lstlisting}

@@

\section{Orchestration of contract and assertions}\label{sec:orchestration}
So far, the assertion and parent contract have been considered separately. However, they have to be orchestrated on the blockchain, such that the assertion code is either invoked automatically before the parent code is executed, or the assertion code can be selectively invoked by the validators. Furthermore, the orchestration should include a mechanism to determine and trigger the required number of test runs per validator. Depending on the orchestration strategy, the compiler then assembles the assertion and parent code into one or several Smart Contracts.

@@

\subsection{Monolithic orchestration}\label{sec:monolithic}
In the monolithic approach, the assertion and parent code are merged into a single Smart Contract. This tightly couples and unites the code in one self-contained unit. As shown in \figref{fig:mono_eps}, the assertions are appended to the original contract as separate entrypoints. Since the \texttt{random} instruction is only available in the execution mode for assertion checking, the assertion code cannot be directly prepended to the respective production code. Otherwise, transactions invoking such an entrypoint could not be included into blocks, as these are validated in the normal execution mode.
\begin{figure}[h]
\centering
  \includegraphics[width=0.5\textwidth]{figures/5-offline_tezos/pipeline_output_mono_ep_basic.jpg}
	\caption{Monolithic assembly - appending assertions as separate entrypoints}
	\label{fig:mono_eps}
\end{figure}

@@

The assembly in \figref{fig:mono_eps}, however, does not yet provide any means to trigger several test runs per validator. Since the required number of test runs $t$ can only be determined at runtime, it cannot be passed to the validators as a parameter to the validation request. Instead, the contract itself should contain a controller to handle the calculation of $t$ and the initiation of as many test runs. Similar to the assertion code, such a controller can be appended as another separate entrypoint, as shown in \figref{fig:monolithic_orchestration}. Instead of invoking the assertion entrypoint (\texttt{A\_a}), the validators would then invoke the controller entrypoint (\texttt{A\_c}), which determines $t$ and invokes (\texttt{A\_a}) $t$ times as an internal operation.
\begin{figure}[h]
\centering
  \includegraphics[width=0.5\textwidth]{figures/5-offline_tezos/pipeline_output_mono_ep}
	\caption{Monolithic assembly with separate entrypoints for controller and assertion}
	\label{fig:monolithic_orchestration}
\end{figure}

As a hybrid solution between a monolithic and a modular approach, the controller can be implemented as a separate Smart Contract, which operates as a proxy. This contract is invoked in place of the monolithic contract, which in turn invokes the original or assertion entrypoints as internal operations. The implementation of the modular part can be adopted from \secref{sec:modular}.

@@

\figref{fig:interaction_monolithic} depicts the resulting interactions of the nodes, validators and monolithic Smart Contracts on the blockchain. The execution model requires a dedicated modus for distributed assertion checking. A possible mechanism to activate this modus is using a special transaction type, which is denoted by \texttt{txa(destination, entrypoint, parameter)} in the graphic. When an operation, such as a transaction or origination, is injected into a node, it is normally propagated to the network after it has been pre-validated by the node. After the injection of \texttt{txa(1, \%A, parameter)}, however, the node broadcasts its derivation \texttt{txa(1, \%A\_c, parameter)} to the network, which invokes the controller entrypoint instead. If this operation finishes successfully for all validators, hence no counterexample is published on the network after a certain amount of block cycles, the node can finally broadcast the initial operation as a normal transaction to the network. In case of an assertion failure, the respective validator broadcasts the counterexample with the generated value to the network. As a consequence of the broadcasted counterexample, the initial operation \texttt{txa(1, \%A, parameter)} should be reverted and rejected by the injecting node. How a counterexample is represented and included on the chain is an unresolved question --- \secref{sec:counterexample} picks this topic up and gives some ideas regarding possible implementations.

@@

\subsubsection{Evaluation of the monolithic approach}
The presented implementations of the monolithic approach each have strengths and weaknesses. Generally, they share a pivotal disadvantage - they require a modification of the parent contract. This complicates the compilation process substantially and precludes the integration of assertions to contracts which have already been deployed to the blockchain. \lstref{lst:mono} shows how the parameter type and branching within the code body have to be adapted compared to the original parent contract in \lstref{lst:mono_parent}. Since the monolithic assembly results in deeper parameter and code nestings, the overall complexity of the contract increases and may obscure the production code.
\lstinputlisting[label=lst:mono_parent, language=Michelson, caption=Skeleton of the pure parent contract in Michelson, numbers=left]{listings/monolithic_parent.tz}

>>>>

As a preliminary and preparation for the compiler implementation, \secref{sec:ext_michelson} identifies necessary extensions to Michelson and Tezos' VM, in order to facilitate the distributed assertion scheme. Although the extensions are formalized in this thesis, implementing the extension is part of the protocol amendment. Furthermore, the compiler needs to generate the target code according to an on-chain orchestration scheme between the assertion and parent contract. This orchestration includes the mechanism to trigger several assertion checks per validator based on \eqref{eq:test_runs}, which was derived in the previous chapter. \secref{sec:orchestration} explores possible strategies for such a scheme, and discusses their strengths and weaknesses. Based on a resulting proposal for the orchestration, \secref{sec:cost_analysis_distributed} performs a cost analysis of distributed assertion checking and evaluates if, and under which conditions, it can increase the scalability in blockchain applications.

As a foundation for the other sections, \secref{sec:tezos} provides important background on the Tezos blockchain, its built-in smart contract language Michelson and some relevant tools and projects within its ecosystem.

@@

In addition to user accounts associated with a public key, Tezos supports smart contracts, which are written in the built-in language Michelson. Tezos' transaction fee system is similar to Ethereum's \cite{wood_ethereum_2021} - it is gas-instrumented and besides a base fee, every operation and byte of storage during contract execution has to be paid for by the user. However, Tezos imposes a hard cap on the amount of gas that can be consumed per operation (including internal transactions) \cite{tezos_docs}\cite{morley_repo}, whereas Ethereum limits the gas quota only in respect to blocks \cite{wood_ethereum_2021}.\\
Tezos is written in the multi-paradigm programming language OCaml. Compiling its source code yields five essential binaries: the  node, baker, endorser, accuser and client. The node is the entity connecting to the peer-to-peer network and keeping a copy of the chain. Bakers are responsible for producing new blocks, the endorsers for validating new blocks and the accusers to call out bakers or endorsers which double-sign or -endorse. The client provides a command line interface to interact with nodes through remote procedure calls (RPC).

\subsection{Proof-of-stake in Tezos}
In Tezos, contracts which have staked a minimum amount of tokens (called a roll), can participate in the consensus mechanism. When participating, a contract has a chance to obtain the role of either a baker or an endorser. Contracts not owning enough tokens or infrastructure to participate directly, can delegate their baking and endorsing rights to other contracts. The rights to bake or endorse are determined and assigned at the beginning of each cycle, which consists of a specified number of blocks. For baking, a random roll is selected for each block level and the rights are assigned to its owner. The block produced by that baker is signed by a fixed number of endorsers\footnote{For consistency, endorsers are referred to as ``validators'' in the remainder of this chapter} (32 as of protocol 007 Delphi), which also have been assigned endorsing rights for this block level by a random selection of rolls. Since participants can stake more than one roll, they may be assigned several endorsement slots at the same block level.

@@

The concrete syntax of Michelson is called Micheline. A program is represented as Micheline nodes, which can be one of the following constructs:
\begin{itemize}
\item A constant of type integer (in decimal notation)
\item A constant of type string
\item A byte sequence in hexadecimal notation
\item An application of a language primitive to a sequence of nodes
\item A sequence of nodes
\end{itemize}
For documentation, readability and additional type constraints, Michelson and Micheline also offer three types of annotations - type, variable and field or constructor annotations, which are labelled with a unique special character in Micheline. The toplevel structure of a smart contract consists of a sequence of the three primitives \texttt{parameter, storage} and \texttt{code}. They declare the type of the input parameter, the storage type and the script, which is interpreted when the contract is invoked. The full grammar of Micheline and Michelson can be found in the Tezos developer resources \cite{tezos_docs}.

\subsubsection{Entrypoints}
Unlike Ethereum's contract language Solidity, Michelson doesn't have a concept of named functions with a dedicated parameter type. Instead of named functions, Michelson contracts model separate entrypoints by taking a sum type as an input parameter. The type constructors can optionally be tagged with a field annotation, which can be considered  as an entrypoint name. The sum type is built by nesting the \texttt{or} data type, which has the constructors \texttt{left} and \texttt{right}. As an example, the contract in \lstref{lst:entrypoints} has two entrypoints --- entrypoint \texttt{A}, which expects an integer wrapped in the \texttt{left} constructor, e.g. \texttt{left 1}, and entrypoint \texttt{B}, which expects a string wrapped in the \texttt{right} constructor, e.g. \texttt{right "hello world"}.
\begin{lstlisting}[language=Michelson, numbers=none, caption=Michelson contract with two entrypoints, label=lst:entrypoints]
parameter (or (int %A) (string %B));
storage unit;
code {
  ...
  IF_LEFT { ... (* entrypoint A *)}
          { ... (* entrypoint B *)}
}
\end{lstlisting}
By adding the constructor annotations \texttt{A} and \texttt{B}, the entrypoints can be invoked explicitly. The passed parameter value is then wrapped automatically into the respective constructor. Furthermore, it is possible to declare a \texttt{default} entrypoint, which is invoked when no explicit tag is specified in the transaction. By default, the default entrypoint is assigned to the root of the parameter type.

\subsubsection{Internal Operations}\label{sec:internal_ops}
Smart contracts are able to invoke other contracts by emitting internal operations. The return type of Michelson programs (more specifically, the type of the top element on the stack after the completed execution) is a tuple of a list of internal operations and the updated storage. After a contract completes, the operations in the returned list are run in sequence. Since internal operations may in turn also emit internal operations, the interpreter puts all emitted operations into a queue, which is processed in order. As an example, \figref{fig:internal_ops} shows an execution order of two external transactions and their internal operations in sequence.
\begin{figure}[h]
\centering
\includegraphics[width=0.5\linewidth]{figures/5-offline_tezos/internal_ops}
\captionsource{Execution order of two external operations and their internal operations}{\cite{tezos_docs}}
\label{fig:internal_ops}
\end{figure}

@@

\subsection{Gas Model}
Tezos stores and transmits data as byte sequences --- in order to obtain a typed representation of their value for interpretation, a byte sequence is first deserialised into Micheline and subsequently parsed into a typed AST. Conversely, before transmitting or storing something on the blockchain, the typed representation is first unparsed into Micheline and then serialised. Since these working steps also entail computational effort, they consume gas and thus account for some of the transaction fees. The gas consumption of an operation (e.g., a transaction or a contract origination) is thus composed of the following positions on top of a base gas fee \cite{morley_gasmodel}\cite{tezos_repo}:
\begin{enumerate}
\item \textbf{Reading costs} --- apply for reading the contract code and storage from the blockchain and depend on the amount of bytes read
\item \textbf{Deserialization costs} --- depend on the components of the Micheline expression and the size of the byte sequence. Apply lazily, thus the deserialization of each value only has to be paid for once
\item \textbf{Parsing costs} --- are composed of three types of costs:
	\begin{itemize}
	\item type parsing --- apply when a Micheline node is converted into a valid type
	\item data parsing --- apply when a node is converted into the value of a known type and depends on the size of the value
	\item code parsing --- apply when the instructions of a program are type checked and depends on the number and types of instructions. Some instructions are expensive, such as \texttt{IF} (both branches need to be checked and compared) or \texttt{CREATE\_CONTRACT} (requires type check of the invoked contract)
	\end{itemize}
\item \textbf{Type comparison costs} --- apply when types are checked for equality (e.g. when type checking the branches of the \texttt{IF} instruction) or during contract interpretation
\item \textbf{Interpretation costs} --- apply for each interpreted instruction and depend on their computational complexity.
\item \textbf{Unparsing costs} --- apply when converting typed data to untyped Micheline nodes and depend on the data type and size
\item \textbf{Serialization costs} --- apply when serializing Micheline nodes to byte sequences and depend on the expression and its size
\item \textbf{Writing costs} --- apply when writing to the internal database and depend on the amount of bytes. A base gas fee is burned for each additional byte of storage.
\end{enumerate}
If a transaction emits internal operations, their gas consumption is computed in the same way and is added to the total gas consumption of the external operation.

\subsection{Developer Tools in the Tezos Ecosystem}
This section describes some tools and projects within Tezos' ecosystem, which are relevant to the development of the toolchain or the protocol amendment.

@@

\subsubsection{Testing Tools}
Before a new protocol is proposed to the Tezos network, it has to be thoroughly tested with system, integration and regression tests. This requires a sandboxed network to simulate the real peer-to-peer network. Tezos' development environment provides the two testing frameworks Flextesa (Flexible network sandboxes)\cite{tezos_docs} and its successor Tezt \cite{tezos_docs}. They allow to configure and run a small, fully functional sandboxed test network including nodes, bakers, endorsers and accusers. Besides testing new protocols, they can also be used to interactively test smart contracts.

\subsubsection{High-level Languages}\label{sec:languages}
Michelson is a compilation target for various high-level languages that provide a more user-friendly and intuitive way of writing smart contracts. Additionally, some of them provide development environments, testing or verification tools. The following list introduces three prominent languages, which compile to Michelson:
\begin{description}
\item[\textbf{Liquidity}] \cite{liquidity} is a language with an OCaml-like syntax, which allows to express contracts in a functional way. It supports using local variables instead of stack manipulations. Its module system can be used to write reusable contract code or libraries. Besides an optimizing compiler, the project also includes a decompiler to compile Michelson programs to Liquidity source code.
\item[\textbf{SmartPy}] \cite{smartpy} is a language available through a Python library and lets developers write contracts and tests using Python syntax and structures (such as classes). Its developer suite includes, i.a., a compiler, a simulation engine for testing contracts and an online editor.
\item [\textbf{Ligo}] \cite{ligo} offers multiple syntax flavours close to Pascal, OCaml or Reason. Besides compilation, the toolchain allows to invoke or evaluate the generated code locally. The project provides Ligo as OCaml packages, which can be integrated into other projects.
\end{description}

@@

\subsection{Language Semantics and Type System}\label{sec:michelson_semantics}
Michelson is interpreted purely functional; the interpreter takes the current stack and an operation and builds a return stack from the initial one, without causing any side effects. The definition of the recursive interpreter is given in the form of a list of rules, comprising of all possible inputs (i.e., program and stack types) and the respective output stack type of the computation. Each rule is of the following form: 
\begin{lstlisting}[caption=Selection rules in the Michelson interpreter \cite{tezos_docs}, language=, numbers=none, label=lst:rules]
> (syntax pattern) / (initial stack pattern)  =>  (result stack pattern)
    iff (conditions)
    where (recursions)
    and (more recursions)
\end{lstlisting}
For each valid program and initial stack, exactly one rule applies. Following the keyword \texttt{iff}, the rule can add extra conditions over values on the stack. If the result depends on the results of other program interpretations, such as conditionals or function calls, the rule can contain recursive rules in the \texttt{where} or \texttt{and} clauses. These clauses represent a recursive interpretation of an intermediate program. The rule thus only applies, if the interpretation of the intermediate result matches the expected pattern.

@@

\subsection{Random Instruction}\label{sec:random}
Since smart contracts generally need to be deterministic, s.t. the network can reach a consensus about the state of the ledger \cite{chatterjee_probabilistic_2019}, the prevalent workarounds for generating pseudo-random numbers cannot be used to implement the random generators in assertion contracts. The scheme explicitly requires the validators to check distinct elements in the search space. Common schemes for obtaining random values, like oracles or using block attributes as seeds \cite{chatterjee_probabilistic_2019}, would generate the same values for all validators. Thus, the VM needs to provide an instruction to generate a ``real'' random value exclusively for assertion checking. In order to preserve the deterministic behaviour of normal smart contracts, the instruction should not be available in the normal execution mode.

\lstref{lst:rand_type} specifies the typing and selection rule of a new instruction \texttt{random} for integers, although such an instruction could also be provided for other data types, such as \texttt{nat}, \texttt{string} or \texttt{mutez}. The selection rule for this instruction applies for the instruction identifier \texttt{RANDOM} and an input stack containing at least two elements. It consumes an offset and a positive range from the stack and pushes a randomly generated integer to the top. The typing rule (line 1) specifies, that the offset should be of type \texttt{integer}, and the range of type \texttt{nat} (i.e., a natural number).
\lstset{upquote=true}
\begin{lstlisting}[caption=Typing and selection rules of the integer \texttt{random} instruction, language=, showlines=true, label=lst:rand_type]
:: int : nat: 'A -> int : 'A
> RANDOM / offset : range : S  => int : S
\end{lstlisting}

Within the Michelson interpreter, the selection of rules is implemented as a huge match case \cite{tezos_repo}. If the rule for \texttt{random} applies, the instruction could be computed in OCaml as indicated in \lstref{lst:rand_impl}.
\begin{lstlisting}[caption=Simplified evaluation of \texttt{random} in the Michelson interpreter, language=, label=lst:rand_impl, float]
match (instruction, stack) with
...
| (Random (offset, (range, rest))) ->
   Random.self_init;
   let rand_int = offset + Random.int(range) in
   return (rand_int, rest)  (* resulting stack *)
\end{lstlisting}

\subsection{List Indexing}\label{sec:nth}
Michelson currently does not provide any predefined indexed access operators for data types like lists, strings or bytes. Accessing a random element of a sequence can, of course, be implemented using low-level instructions. However, because such operations are used frequently in many of the given use-cases, adding it as a predefined primitive for some data types might be justifiable. It reduces the required code to express the indexing, and thus not only improves the readability of the contract code, but also lowers the costs of the contract origination and execution.

\lstref{lst:nth_type} specifies the typing and selection rules of an instruction \texttt{nth} to access the element of a list at a specified index. It consumes a list with elements of type \texttt{'a}, and an index of type \texttt{nat} from the top of the stack. After the computation, on top of the resulting stack is an optional value of the element type \texttt{'a}. A selection rule (line 2 and 3) is given for each of the two possible result stack patterns, i.e., a value of type \texttt{'a} or an undefined value if the given index is out of bounds.
\begin{lstlisting}[caption=Selection and type rule of the list \texttt{nth} instruction, language=, label=lst:nth_type]
:: (list 'a) : nat : 'A -> option 'a : 'A
> NTH / l : index : S  =>  Some 'a: S
     iff index is within bounds
> NTH / l : index : S  =>  None 'a: S
     iff index is out of bounds
\end{lstlisting}

@@

\section{Orchestration of Contract and Assertions}\label{sec:orchestration}
So far, the assertion and parent contract have been considered separately. However, they have to be orchestrated on the blockchain, such that the assertion code is either invoked automatically before the parent code is executed, or the assertion code can be selectively invoked by the validators. Furthermore, the orchestration should include a mechanism to determine and trigger the required number of test runs per validator. Depending on the orchestration strategy, the compiler then assembles and translates the assertion and parent code into one or several smart contracts.

@@

\subsection{Monolithic Orchestration}\label{sec:monolithic}
In the monolithic approach, the assertion and parent code are merged into a single smart contract. This tightly couples and unites the code in one self-contained unit. As shown in \figref{fig:mono_eps}, the assertions are appended to the original contract as separate entrypoints. Since the \texttt{random} instruction is only available in the execution mode for assertion checking, the assertion code cannot be directly prepended to the respective production code. Otherwise, transactions invoking such an entrypoint could not be included into blocks, as these are validated in the normal execution mode.
\begin{figure}[h]
\centering
  \includegraphics[width=0.5\textwidth]{figures/5-offline_tezos/pipeline_output_mono_ep_basic.jpg}
	\caption{Monolithic assembly - appending assertions as separate entrypoints}
	\label{fig:mono_eps}
\end{figure}

@@

The assembly in \figref{fig:mono_eps}, however, does not yet provide any means to trigger several test runs per validator. Since the required number of test runs $t$ can only be determined at runtime, it cannot be passed to the validators as a parameter to the validation request. Instead, the contract itself should contain a controller to handle the calculation of $t$ and the initiation of $t$ test runs. Similar to the assertion code, such a controller can be appended as another separate entrypoint, as shown in \figref{fig:monolithic_orchestration}. Instead of invoking the assertion entrypoint (\texttt{A\_a}), the validators would then invoke the controller entrypoint (\texttt{A\_c}), which determines $t$ and invokes (\texttt{A\_a}) $t$ times as an internal operation.
\begin{figure}[h]
\centering
  \includegraphics[width=0.5\textwidth]{figures/5-offline_tezos/pipeline_output_mono_ep}
	\caption{Monolithic assembly with separate entrypoints for controller and assertion}
	\label{fig:monolithic_orchestration}
\end{figure}

As a hybrid solution between a monolithic and a modular approach, the controller can be implemented as a separate smart contract, which operates as a proxy. This contract is invoked in place of the monolithic contract, which in turn invokes the original or assertion entrypoints as internal operations. The implementation of the modular part can be adopted from \secref{sec:modular}.

@@

\figref{fig:interaction_monolithic} depicts the resulting interactions of the nodes, validators and monolithic smart contracts on the blockchain. The execution model requires a dedicated modus for distributed assertion checking. A possible mechanism to activate this modus is using a special transaction type, which is denoted by \texttt{txa(destination, entrypoint, parameter)} in the graphic. When an operation, such as a transaction or origination, is injected into a node, it is normally propagated to the network after it has been pre-validated by the node. After the injection of \texttt{txa(1, \%A, parameter)}, however, the node broadcasts its derivation \texttt{txa(1, \%A\_c, parameter)} to the network, which invokes the controller entrypoint instead. If this operation finishes successfully for all validators, hence no counterexample is published on the network after a certain amount of block cycles, the node can finally broadcast the initial operation as a normal transaction to the network. In case of an assertion failure, the respective validator broadcasts the counterexample with the generated value to the network. As a consequence of the broadcasted counterexample, the initial operation \texttt{txa(1, \%A, parameter)} should be reverted and rejected by the injecting node. How a counterexample is represented and included on the chain is an unresolved question --- \secref{sec:counterexample} picks this topic up and gives some ideas regarding possible implementations.

@@

\subsubsection{Evaluation of the Monolithic Approach}
The presented implementations of the monolithic approach each have strengths and weaknesses. Generally, they share a pivotal disadvantage - they require a modification of the parent contract. This complicates the compilation process substantially and precludes the integration of assertions to contracts which have already been deployed to the blockchain. \lstref{lst:mono} shows how the parameter type and branching within the code body have to be adapted compared to the original parent contract in \lstref{lst:mono_parent}. Since the monolithic assembly results in deeper parameter and code nestings, the overall complexity of the contract increases and may obscure the production code.
\lstinputlisting[label=lst:mono_parent, language=Michelson, caption=Skeleton of the pure parent contract in Michelson, numbers=left]{listings/monolithic_parent.tz}
